---
title: "California Wine  - Part 1"
author: "Daniel Mikel"
date: "2017-11-28T21:13:14-05:00"
output: html_document
runtime: shiny
---

The [United States Department of Agriculture](https://www.nass.usda.gov/Statistics_by_State/California/Publications/Grape_Crush/Reports/) publishes contract level price and volume data for California grapes crushed and used for wine. This allows us to see each purchase of a wine variety, what variety, it's price, and the tonnage in the purchased amoung, giving us a disaggregated level data of data ripe for analysis. This is intended to be the first in a series of blog posts analysing the dynamics of the California.

```{r, message = FALSE, echo = FALSE}
library(ggridges)
library(ggjoy)
library(tidyverse)
library(knitr)

library(shiny)
library(shinythemes)
library(scales)

gc <- read.csv("/home/dan/Data/NASS/Grape_Crush/gc_csv/all_gc_clean.csv")
gc$District_Number <- as.factor(gc$District_Number)

var <- gc %>%
  filter(Variety != "all red",
         Variety != "all white",
         Variety != "all wine")
types <- gc %>%
  filter(Variety == "all red" |
           Variety == "all white" |
           Variety == "all wine")

cols <- c("red" = "#F8766D", "white" = "#00BFC4", "raisin" = "#C77CFF", "table" = "#7CAE00", "all wine" = "#190B28")

Simple_Cap <- function(x) {
  s <- strsplit(x, " ")[[1]]
  paste(toupper(substring(s, 1,1)), substring(s, 2),
        sep="", collapse=" ")
}

fun_title <- function(variety){
  if(length(variety) >= 3){
    paste0(str_c(sapply((variety[1:(length(variety)-1)]), Simple_Cap), collapse = ", "), " and ", sapply((variety[length(variety)]), Simple_Cap))
  } else if (length(variety) == 1){
    paste0(sapply((variety), Simple_Cap))
  } else {
    str_c(sapply((variety), Simple_Cap), collapse = " and ")
  }
}
```

```{r, echo = FALSE, results = 'asis'}
knitr::kable(gc[1:5, 1:6])
```

We can see the distribution of tonnage within California's wine regions.

```{r, echo = FALSE, message = FALSE, fig.height = 6}
cols <- c("red" = "#F8766D", "white" = "#00BFC4", "raisin" = "#C77CFF", "table" = "#7CAE00", "all wine" = "#190B28")

gc %>%
  filter(Year == 2016) %>%
  group_by(District_Disc, District_Number, Type) %>%
  summarise(Thousand_Tons = sum(Tons)/1000) %>%
  arrange(District_Number) %>%
  ggplot(aes(x = as.factor(District_Number), y = Thousand_Tons, fill = Type)) +
  geom_col() + 
  scale_fill_manual(values = cols) +
  xlab(NULL) + 
  ylab("Thousand Tons Sold") + 
  scale_x_discrete(labels = c("1" = "[1] Mendocino",
                              "2" = "[2] Lake",
                              "3" = "[3] Sonoma",
                              "4" = "[4] Napa",
                              "5" = "[5] Solano",
                              "6" = "[6] North Central Coast",
                              "7" = "[7] Monterey and San Benito",
                              "8" = "[8] San Luis Obispo",
                              "9" = "[9] Northern California",
                              "10" = "[10] Sierra Foothills",
                              "11" = "[11] San Joaquin and Sacramento",
                              "12" = "[12] San Joaquin, Stanislaus, Merced",
                              "13" = "[13] Fresno, Madera, N. Tulare, Mono, Inyo",
                              "14" = "[14] Kern, Kings, S. Tulare",
                              "15" = "[15] L.A. and San Bernardino",
                              "16" = "[16] So Cal",
                              "17" = "[17] Yolo")) +
  ggtitle("Grape Production in Tons by District") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))  
```

And can see which districts grow which varieties

```{r, echo = FALSE, fig.height = 9}
library(plotly)
library(shiny)

ui <- fluidPage(
  sliderInput("Year", "Choose a year:", 
            min = 2005, max = 2016, 
            value = 2017,
            sep = ""),
selectizeInput("variety", "Select a variety and year",
               sort(unique(gc$Variety)),
               selected = c("all red"),
               multiple = FALSE),
radioButtons("Vort", "Display Production in ",
             c("Tons" = "Tons",
               "Dollars" = "Dollars")),
  plotlyOutput("plot")
)

server <- function(input, output, session) {

  output$plot <- renderPlotly({
    # use the key aesthetic/argument to help uniquely identify selected observations
    if (input$Vort == "Tons"){
      p <- gc %>%
      filter(Year == input$Year,
             Variety == input$variety) %>%
      group_by(District_Disc, District_Number, Type) %>%
      summarise(Yaxis = sum(Tons)) %>%
      arrange(District_Number) %>%
      ggplot(aes(x = as.factor(District_Number), y = Yaxis, fill = Type)) +
      geom_col() + 
      scale_fill_manual(values = cols) +
      xlab(NULL) + 
      ylab(paste("Production in", input$Vort)) + 
      scale_x_discrete(labels = c("1" = "[1] Mendocino",
                                  "2" = "[2] Lake",
                                  "3" = "[3] Sonoma",
                                  "4" = "[4] Napa",
                                  "5" = "[5] Solano",
                                  "6" = "[6] North Central Coast",
                                  "7" = "[7] Monterey and San Benito",
                                  "8" = "[8] San Luis Obispo",
                                  "9" = "[9] Northern California",
                                  "10" = "[10] Sierra Foothills",
                                  "11" = "[11] San Joaquin and Sacramento",
                                  "12" = "[12] San Joaquin, Stanislaus, Merced",
                                  "13" = "[13] Fresno, Madera, N. Tulare, Mono, Inyo",
                                  "14" = "[14] Kern, Kings, S. Tulare",
                                  "15" = "[15] L.A. and San Bernardino",
                                  "16" = "[16] So Cal",
                                  "17" = "[17] Yolo")) +
     scale_y_continuous(labels = comma) +
     ggtitle(paste("Production in", input$Vort, "for", fun_title(input$variety), "Wine Grapes by District")) +
     theme_minimal() +
     theme(axis.text.x = element_text(angle = 30, hjust = 1),
           plot.title = element_text(size = 18))
      ggplotly(p)
    } else {
      p <- gc %>%
      filter(Year == input$Year,
             Variety == input$variety) %>%
      group_by(District_Disc, District_Number, Type) %>%
      summarise(Yaxis = sum(Tons * Base.Price.Per.Ton)) %>%
      arrange(District_Number) %>%
      ggplot(aes(x = as.factor(District_Number), y = Yaxis, fill = Type)) +
      geom_col() + 
      scale_fill_manual(values = cols) +
      xlab(NULL) + 
      ylab(paste("Production in", input$Vort)) + 
      scale_x_discrete(labels = c("1" = "[1] Mendocino",
                                  "2" = "[2] Lake",
                                  "3" = "[3] Sonoma",
                                  "4" = "[4] Napa",
                                  "5" = "[5] Solano",
                                  "6" = "[6] North Central Coast",
                                  "7" = "[7] Monterey and San Benito",
                                  "8" = "[8] San Luis Obispo",
                                  "9" = "[9] Northern California",
                                  "10" = "[10] Sierra Foothills",
                                  "11" = "[11] San Joaquin and Sacramento",
                                  "12" = "[12] San Joaquin, Stanislaus, Merced",
                                  "13" = "[13] Fresno, Madera, N. Tulare, Mono, Inyo",
                                  "14" = "[14] Kern, Kings, S. Tulare",
                                  "15" = "[15] L.A. and San Bernardino",
                                  "16" = "[16] So Cal",
                                  "17" = "[17] Yolo")) +
     scale_y_continuous(labels = comma) +
     ggtitle(paste("Production in", input$Vort, "for", fun_title(input$variety), "Wine Grapes by District")) +
     theme_minimal() +
     theme(axis.text.x = element_text(angle = 30, hjust = 1),
           plot.title = element_text(size = 18))
      ggplotly(p)
    }
  })
}

shinyApp(ui, server)

```


We can also see a distribution of prices by district. We can see a sharp difference between the higher production regions like Fresno, Madera, and North Tulare counties and the districts average weighted price. 

```{r, echo = FALSE, message = FALSE, fig.height = 6}
gc %>%
  filter(Base.Price.Per.Ton <= 10000) %>%
  arrange(desc(District_Number)) %>%
  mutate(label = factor(District_Disc, unique(District_Disc))) %>%
  ggplot(aes(x = Base.Price.Per.Ton, y = label, fill = label)) + 
  geom_joy(rel_min_height = 0.005, scale = 2,alpha = 0.5) +
  # scale_x_continuous(breaks = c(0,5000, 10000, 15000)) +
  scale_y_discrete(expand = c(0.01, 0)) +
  scale_fill_cyclical(values = c("#66c2a5",
                                 "#fc8d62",
                                 "#8da0cb",
                                 "#e78ac3",
                                 "#a6d854",
                                 "#ffd92f",
                                 "#e5c494")) +
  ggtitle("Histograms of Grape Prices by District") +
  theme_joy() +
  theme(axis.text = element_text(size = 10),
        axis.text.y = element_text(size = 8, angle = 0)) +
  xlab("Price, Dollars per Ton") +
  ylab(NULL) 
```

Price between varieties seems pretty coorolated within districts between varieties, with red wine consistently outpricing white wine in each district

```{r, echo = FALSE, message = FALSE, fig.height = 6}
gc %>%
  group_by(District_Disc, District_Number, Year, Type) %>%
  filter(Type %in% c("red", "white")) %>%
  summarize(Average_Price = sum(Base.Price.Per.Ton * Tons)/sum(Tons))  %>%
  spread(Type, Average_Price) %>%
  ggplot(aes(x = red, y = white)) +
  geom_point() + 
  geom_text(aes(label = District_Number), hjust = 1, vjust = 1.5) +
  geom_smooth(aes(group = 1), method = lm, se = FALSE, linetype = 2) +
  ggtitle("Red and White Grower Return Prices, Weighted Dollars per Ton") +
  theme_minimal()
```

We can also see that over time, as prices fluctuate over the years, the relationship between white and red wine prices stays consistent. Here we track the price in Sonoma County from 2005 to the present. We can see that while prices have changed considerably year over year, red wine has stayed abour 1.4x the price of white wine within the district

```{r, echo = FALSE, message = FALSE, fig.height = 6}
gc %>%
  group_by(District_Disc, District_Number, Year, Type) %>%
  filter(Type %in% c("red", "white")) %>%
  summarize(Average_Price = sum(Base.Price.Per.Ton * Tons)/sum(Tons))  %>%
  spread(Type, Average_Price) %>%
  filter(District_Number == 3) %>%
  ggplot(aes(x = red, y = white)) +
  geom_point() + 
  geom_text(aes(label = Year), hjust = 1, vjust = 1.5) +
  geom_smooth(aes(group = 1), method = lm, se = FALSE, linetype = 2) +
  ggtitle("Historic Red and White Grower Return Prices for Napa County, Weighted Dollars per Ton") +
  theme_minimal()     
```

This  can be seen across many California Wine Districts.

```{r, echo = FALSE, message = FALSE, fig.height = 6}
gc %>%
  group_by(District_Disc, District_Number, Year, Type) %>%
  filter(Type %in% c("red", "white")) %>%
  summarize(Average_Price = sum(Base.Price.Per.Ton * Tons)/sum(Tons))  %>%
  spread(Type, Average_Price) %>%
  filter(District_Number %in% c(1:4, 16)) %>%
  ggplot(aes(x = red, y = white, color = District_Number)) +
  geom_point() + 
  # geom_text(aes(label = Year), hjust = 1, vjust = 1.5) +
  geom_smooth(method = lm, se = FALSE, linetype = 2) +
  ggtitle("Historic Red and White Grower Return Prices By District, Dollars per Ton") +
  theme_minimal()     
```

```{r, echo = FALSE, message = FALSE, fig.height = 6}
gc %>%
  group_by(District_Disc, District_Number, Year, Type) %>%
  filter(Type %in% c("red", "white")) %>%
  summarize(Average_Price = sum(Base.Price.Per.Ton * Tons)/sum(Tons))  %>%
  spread(Type, Average_Price) %>%
  filter(District_Number %in% c(5, 7,9, 11:14, 17)) %>%
  ggplot(aes(x = red, y = white, color = District_Number)) +
  geom_point() + 
  # geom_text(aes(label = Year), hjust = 1, vjust = 1.5) +
  geom_smooth(method = lm, se = FALSE, linetype = 2) +
  ggtitle("Historic Red and White Grower Return Prices By District, Dollars per Ton") +
  theme_minimal()  
```

But is not necessarily present in every district
```{r, echo = FALSE, message = FALSE, fig.height = 6}
gc %>%
  group_by(District_Disc, District_Number, Year, Type) %>%
  filter(Type %in% c("red", "white")) %>%
  summarize(Average_Price = sum(Base.Price.Per.Ton * Tons)/sum(Tons))  %>%
  spread(Type, Average_Price) %>%
  filter(District_Number %in% c(6, 8, 10, 15)) %>%
  ggplot(aes(x = red, y = white, color = District_Number)) +
  geom_point() + 
  # geom_text(aes(label = Year), hjust = 1, vjust = 1.5) +
  geom_smooth(method = lm, se = FALSE, linetype = 2) +
  ggtitle("Historic Red and White Grower Return Prices By District, Dollars per Ton") +
  theme_minimal()     
```

We can compare historic varieties prices within regions. The interactive graph below allows you to select varieties or types within a region and compare it to other varieties and types within the same region. A full version can be found [daniel-mikel.shinyapps.io](https://daniel-mikel.shinyapps.io/historic_california_grape_crush/)

```{r, echo = FALSE, fig.width = 9}

knitr::include_app("https://daniel-mikel.shinyapps.io/gc_variety_embed/", 
                   height = "650px")

```